Topological Equivariant Artist Model for Visualization Library Architecture

### 2023-06-09 paper/main.tex
numberline

% theorem environments
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
\newtheorem{axiom}{Axiom}
\newtheorem{prop}{Proposition}

% *** diagrams
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\usepackage{pgfplots} %numberline
\DeclareUnicodeCharacter{2212}{−}
\usepgfplotslibrary{groupplots,dateplot}
\usetikzlibrary{patterns,shapes.arrows}
\pgfplotsset{compat=newest}

@@

\subsection{Topological Structure}
\label{sec:related-work:continuity}
\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{k_different_types.png}
  \caption{
  \note{add first panel which is table + all K type, second is scatter, etc...}
%
  \label{fig:related-work:continuity:ktypes}}
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
  \caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction. These local trivializations are themselves locally trivial fiber bundles.} 
  \label{fig:atct:local_trivalizations}
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
\end{figure}
As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.

@@

For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
\begin{minipage}[c]{0.25\columnwidth} 
  \begin{equation}
    \label{eq:atct:base_union}
  \begin{tikzcd}
    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
    \dbasec_a \arrow[u, "\iota_a", color=base] 
    & \dbasec_c 
    \arrow[u, "\iota_c"', color=base] 
    \arrow[l, "\iota_c"', color=base] 
  \end{tikzcd}
\end{equation}
\end{minipage}\hfill

\begin{minipage}[c]{0.65\columnwidth}
  \note{will be replaced by lines showing intervals}
  \begin{tikzpicture}[commutative diagrams/every diagram]
    \matrix[matrix of math nodes, name=m, commutative diagrams/every cell]{
      (0,1) & (.55,1) \\
      (0,.45) & (.40, 60)\\  
    };

    \path[commutative diagrams/.cd, every arrow, every label]
    (m-2-2) edge[color=base] (m-1-2)
            edge[color=base] (m-2-1)
    (m-2-1) edge[color=base] (m-1-1)
    (m-1-2) edge[color=base] (m-1-1);
  \end{tikzpicture}
\end{minipage}

@@

\begin{minipage}[c]{.25\columnwidth}
\begin{tikzcd}
  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
  & \dfiberc^{b} \arrow[d, "\pi_c"] 
  & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
  & {(pressure,time)} \arrow[d, "\pi_c"] \\
  \dfiberc^{a} \arrow[r, "\pi_c"] 
  & \dfiberc^{c}
  & {(temperature,time)} \arrow[r, "\pi_c"]
  & time
  \end{tikzcd}
\end{minipage}\hfill
\begin{minipage}[c]{.65\columnwidth}
  \note{will be little pictures like in 7 sketchs}
\end{minipage}

@@

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{dbundle.png}
  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section. Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base space encoding temporal continuity. Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
    \label{fig:atct:trivialbundle}}
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB pre-render space with infinite resolution $\mathcal{R}^{2}$. Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space. \note{add alpha and z channels and very faded out marker point} 
  \label{fig:atct:fb:graphic}}
\end{figure}

@@

We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way. As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, discussed in \autoref{eq:artist:operator}, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).

@@

\subsubsection{Example: Graphic and Data}
\begin{figure}[h!]
  \includegraphics*[width=1\columnwidth]{xi_sin.png}
  \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space. We choose to visualize this as a circle and two line plots. The indexing function \vindex\ book keeps \note{find better word than bookkeeping} which parts of the circle and each curve correspond to each point on the unit circle. The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point. \label{fig:atct:morphisms:sheaf}} 
\end{figure}

@@

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{phi.png}
  \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g. doubled; $\dfunc$ - values can change position and value  \label{fig:atct:phi}}
\end{figure}
\autoref{fig:atct:phi} provides an example of a transposition base space change \dfunch, a scaling fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$. In the transposition only case, the values in $\dfuncpull\dsection$ retain their neighbors from $\dsection$ because \dfunc\ does not change the continuity. Each value in $\dfuncpull\dsection$ is also the same as in $\dsection$, just moved to the new position. In $\dfunct\dsection$, each value is scaled by two but remains in the same location as in $\dsection$. And in $\dfunc\dsection$ each function is transposed such that it retains its neighbors and all values are scaled consistently.

@@

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{equivariance.png}
  \caption{This artist is equivariant because when the input data $\dsection$ is transposed, $\dfunch$, scaled $\dfunct$, and transposed and scaled $\dfunc$, the corresponding colored cells are transposed, scaled such that the color is moved two steps, and both transposed and scaled. 
  \label{fig:artist:equivariance}}
\end{figure}

@@

\begin{equation*}
  \label{eq:artist:multiplication}
  \vartistc_{a \times b}(\cgamma{\dbase}{\dtotalc^{a\times b}}) \coloneqq \vartistc_{a}(\cgamma{\dbasec}{\dtotalc^{a}}) \times \vartistc_{b}(\cgamma{\dbasec}{\dtotalc^{b}}) 
\end{equation*}
which following from an adjoint property of homsets \note{find citation}
\begin{equation*}
  Hom(\gbase, \gfiber) \times Hom(\gbase, \gfiber) = Hom(\gbase, \gfiber\times \gfiber)
\end{equation*}
which means that the artists on the subsets of fibers can be defined 
\begin{equation}
  \gsectionc^{a \times b} = \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
\end{equation} 
but that the signature of $\gsectionc^{a \times b}$ would be $\gbase \rightarrow \gfiber \times \gfiber$. Instead of having to special case the return type of artists that are compositions of multiple case, the hom adjoint \note{find cite} property
\begin{equation*}
  Hom(\gbase, \gfiber \times \gfiber) = Hom(\gbase+\gbase, \gfiber)
\end{equation*}
 means that multiplication can be considered as a special case of addition where $\dbase^{a} = \dbase^{b}$. While we discussed the trivial case in \autoref{sec:artist:addition}, there is no strict  requirement that $\dfiber^{a} = \dfiber^{b}$.

@@

\begin{figure}[h!]
  \centering
  \includegraphics*[scale=.75]{qcom.png}
  \caption{The circle-line visual element can be constructed via $\gsection_{circle}$ + $\gsection_{line}$ functions that generate the circle and line elements respectively. This is equivalent to a $\gsection_{circle+line}$ function that takes as input the combined base space $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$ and returns pixels in the circle-line element.  \label{fig:artist:operator}}
\end{figure}
There is no way to visually determine whether a visual element is the output of a single artist or a multiplied or added collection of artists. The circle-line visual element in \autoref{fig:artist:operator} can be a visual representation of a highlighted point intersecting with a line plot with the same fields. The same element can also be encoding some fields of a section in the circle and other fields of that section in the lines. \note{+*equive}
Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
\begin{equation}
  \vartistc(\cgamma{\mathop{\sqcup}_{i} \dbasec^{i}}{\mathop{\oplus}_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
  \vartistc_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
\end{equation}

As shown in \autoref{eq:artist:operator}, bundles over a union of base spaces can be joined as a product of the fibers. This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \mathop{\sqcup}_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends- 
and therefore which commonalities need to be preserved in the artists that act on these inputs.

@@

\begin{figure}[h!]

@@

\begin{figure}[h!]
  \includegraphics[width=1\columnwidth]{full_scatter.png}
  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q} \note{much better labeling, include semantic labeling, make everything bigger}}
\end{figure}
As shown in \autoref{fig:construction:q}, a set of  $\vchannel$ functions individually convert the values in the data record to visual components. Then the $\vmark$ function combines these visual encodings to produce a graphic section \gsection. When this section is evaluated on the graphic space associated with the data $\gsection(\vindexpre(\dbasepoint))$, it produces a blue circular marker at the intersection of the x and y positions listed in \vsection. The composition rule in \autoref{eq:construction:q:fabrication} means that developers can implement $\vmark$ as drawing circles or can implement a $\vmark$ that draws arbitrary shapes, and then provide different $\vchannel$ adapters, such as one that specifies that the shape is a circle.

@@

\subsection{Bundle Inspired Data Containers}
\begin{table}[h!]
  \centering
\begin{tabular}{|lcl|}
  \hline \\
   fruit &  calories &  juice \\
  \hline\\
    apple &        95 &   True \\ 
   orange &        67 &   True \\ 
  lemon &        17 &  False \\ 
      lime &        20 &  False \\
  \hline
\end{tabular}
\label{tab:discussion:data}
\end{table}
We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$. We thinly wrap \autoref{tab:discussion:data} in an object so that the common data interface function is that $\tau = \texttt{DataContainerObject.query}$. 
\begin{minted}{python}
  class FruitFrameWrapper:
    def query(self, data_bounds, sampling_rate):
      # local sections are a list of
      # {field: local_batch_of_values}
      return local_sections
\end{minted}

This interface provides a uniform way of accessing subsets of the data, which are local sections. The motivation for a common data interface is that it would allow the artist to talk to different common python data containers, such as numpy\cite{harris2020array}, pandas\cite{jeff_reback_2020_3715232}, xarray \cite{hoyer2017xarray}, and networkx\cite{HagbergExploringNetwork2008}. Currently, data stored in these containers must be unpacked and converted into arrays and matrices in ways that either destroy or recreate the structure encoded in the container. For example a pandas data frame must be unpacked into its columns before it is sent into most artists and continuity is implicit in the columns being the same length rather than a tracked base space $\dbase$. Because it is more efficient to work with the data in column order, we often treat the data as a collection of single fiber bundles. This is equivalent to the total bundle, as shown in \autoref{eq:artist:operator}.

@@

\begin{table}[h!]
  \centering
  {\renewcommand{\arraystretch}{1.5}
  \begin{tabular}{|r | c c c|}
    \hline
    &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
     &  location/subset/indices & record/fields &  dataset type\\
    \hline
   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & $\delementc \in \dfiberc$ & \dtotalc\\
   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & $\velementc \in \vfiberc$ & \vtotalc\\
   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & $\gelementc \in \gfiberc$ & \gtotalc\\
   \hline
  \end{tabular}
  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles}}
  \label{tab:appendix:summary:objects}
  }
\end{table}

\begin{table}[h!]
  \centering
  {\renewcommand{\arraystretch}{1.5}
  \begin{tabular}{|r | l l | }
    \hline
     & \textcolor{section}{section} & \textcolor{sheaf}{sheaf} \\
     & record at location & set of possible records for subset \\
     \hline
  Data & $ \cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$\\
  Visual &  $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$\\
  Graphic &    $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ &  $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ \\ 
  \hline
  \end{tabular}
  \caption{Functions that associate topological subspaces with records, discussed in \autoref{sec:atct:fb:sections} and \autoref{sec:atct:sheaves}}
  \label{tab:appendix:summary:datafunctions}
  }
\end{table}

>>>>

% theorem environments
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newcommand{\definitionautorefname}{Definition}
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
\newtheorem{axiom}{Axiom}
\newtheorem{prop}{Proposition}

% *** diagrams
\usepackage{tikz}
\usetikzlibrary{cd, calc, positioning}
\usetikzlibrary{patterns,shapes.arrows}
\usepackage{pgfplots} %numberline
\usepgfplotslibrary{groupplots,dateplot}
\pgfplotsset{
  compat=1.15,
   simple axis/.style={
    scale only axis,
    axis x line=bottom,
    axis y line=none,
    axis line style={latex-latex},
    xticklabel style = {font=\footnotesize},
    width=40pt,
    height=20pt,
    xtick = {0, 180, 360},
    xticklabels = {$0$, $\pi$, $2\pi$},
    xmin = 0, xmax=360,
    ymin=0, ymax=1,
    domain=0:360,
    range=0:1,
    tickwidth=0,
    }
}
\pgfplotsset{compat=newest}

@@

\subsection{Topological Structure}
\label{sec:related-work:continuity}
\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{k_different_types.png}
  \caption{
  \note{add first panel which is table + all K type, second is scatter, etc...}
%
  \label{fig:related-work:continuity:ktypes}}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\columnwidth]{figures/local_trivalizations.png}
  \caption{The local trivializations $\dtotal_{0}$ and $\dtotal_{1}$ are subspaces of $\dtotal$ where the fibers $\dfiber_{0}$ are equal, here illustrated as all pointing in the same direction. These local trivializations are themselves locally trivial fiber bundles.} 
  \label{fig:atct:local_trivalizations}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{figures/cyl_mob.png}
  \caption{\note{figure out short caption here}\label{fig:cyl_mob_bundles}}
\end{figure}
As demonstrated in \autoref{fig:cyl_mob_bundles} different types of total spaces can be decomposed into the same local trivializations; therefore the distinguishing factor between total spaces can be the transition maps for constructing the spaces from trivializations.

@@

For example, the disjoint union of two bundles, as shown in \autoref{fig:atct:local_trivalizations}, is the coproduct $\dbase^{a} \subseteq_{\dbase^{c}} \dbase^{b}$ over an overlap $\dbase^{c}$ and therefore the inclusion morphism must be commutative:
\begin{minipage}[c]{0.30\columnwidth} 
  \begin{equation}
    \label{eq:atct:base_union}
  \begin{tikzcd}
    \dbasec^{a}\sqcup_{\dbasec^{c}}\dbasec^{b} 
    & \dbasec_{b} \arrow[l, "\iota_b"', color=base] \\
    \dbasec_a \arrow[u, "\iota_a", color=base] 
    & \dbasec_c 
    \arrow[u, "\iota_c"', color=base] 
    \arrow[l, "\iota_c"', color=base] 
  \end{tikzcd}
\end{equation}
\end{minipage}\hfill
\begin{minipage}[c]{0.60\columnwidth}
  \begin{tikzpicture}[commutative diagrams/every diagram]
    \draw node[anchor=north west, minimum width=30pt, minimum height=10pt] (kabc) {};
\begin{axis}[simple axis, at={(kabc)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\draw node[anchor=north east, right=55pt of kabc, minimum width=30pt, minimum height=10pt] (kb) {};
%\begin{axis}[simple axis,at={(kb)}]
%  %\addplot [thick, smooth] {sin(2*pi*x)};
%\end{axis}
\draw[at=(kb)][draw=black](0,0) -- (360,0);
\draw node[anchor=south west, below=of kabc, minimum width=30pt, minimum height=10pt] (ka) {};
\begin{axis}[simple axis,at={(ka)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\draw node[anchor=south east, below=of kb, minimum width=30pt, minimum height=10pt] (kc) {};
\begin{axis}[simple axis, at={(kc)}]
  %\addplot [thick, smooth] {sin(2*pi*x)};
\end{axis}

\path[commutative diagrams/.cd, every arrow, every label]
  (kc) edge[] (kb)
       edge[] (ka)
  (ka) edge[] (kabc)
  (kb) edge[] (kabc);

  \end{tikzpicture}

\end{minipage}

@@

\begin{minipage}[c]{.25\columnwidth}
\begin{tikzcd}
  \label{eq:atct:fiber_product}
  \dfiberc^{a}\times_{\dfiber^{c}} \dfiberc^{b} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
  & \dfiberc^{b} \arrow[d, "\pi_c"] 
  & {(temperature, time) \times_{time} (pressure, time)} \arrow[d, "\pi_a"] \arrow[r, "\pi_b"] 
  & {(pressure,time)} \arrow[d, "\pi_c"] \\
  \dfiberc^{a} \arrow[r, "\pi_c"] 
  & \dfiberc^{c}
  & {(temperature,time)} \arrow[r, "\pi_c"]
  & time
  \end{tikzcd}
\end{minipage}\hfill
\begin{minipage}[c]{.65\columnwidth}
  \note{will be little pictures like in 7 sketchs}
\end{minipage}

@@

\begin{figure}[H]
  \includegraphics[width=\columnwidth]{dbundle.png}
  \caption{The table from \autoref{fig:related-work:continuity:ktypes} has a 3 dimensional fiber (name, temperature, precipitation), a 0D base space, and each row is a section. Each time series is a section of a bundle with a 2D fiber (time, precipitation) and a 1D base space encoding temporal continuity. Each location of the rain map is a section of a bundle with a 2D plane base encoding spatial continuity and a 3D fiber space encoding (latitude, longitude, precipitation)
    \label{fig:atct:trivialbundle}}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{fb_rho.png}
  \caption{For a 2D display, a section $\gsection$ maps from each point $\gbasepoint$ into a fiber $\gfiber$ that encodes an RGB pre-render space with infinite resolution $\mathcal{R}^{2}$. Each tiny colored box is an approximation of the return value of the same section function $\gsection$ evaluated on different points $\gbasepoint \in \gbase$ in the base space. \note{add alpha and z channels and very faded out marker point} 
  \label{fig:atct:fb:graphic}}
\end{figure}

@@

We construct the map as going from graphic to data because that encodes the notion that every visual element traces back to the data in some way. As exemplified in \autoref{fig:atct:morphisms:sheaf}, we define $\vindex$ as a surjective map because it allows us to express that a union of graphic spaces $S_i$ maps to single data point $\dbasepoint$, which allows us to express visual representations of a single record that are the union of many primitives, such as multipart glyphs (e.g boxplots) and combinations of plot types (e.g line with point markers).

@@

\subsubsection{Example: Graphic and Data}
\begin{figure}[H]
  \includegraphics*[width=1\columnwidth]{xi_sin.png}
  \caption{The data consists of the $sin$ and $cos$ functions over a unit circle base space. We choose to visualize this as a circle and two line plots. The indexing function \vindex\ book keeps \note{find better word than bookkeeping} which parts of the circle and each curve correspond to each point on the unit circle. The pushforward $\vindexpush$ matches each point in the data space to the specification of the graphic at that point, while the pullback $\vindexpull$ matches each point in the graphic space to the data over that point. \label{fig:atct:morphisms:sheaf}} 
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\columnwidth]{phi.png}
  \caption{Values in a data set can be transformed in three ways: $\dfunch$-values can change position, .e.g transposed;  $\dfunct$-values can change, e.g. doubled; $\dfunc$ - values can change position and value  \label{fig:atct:phi}}
\end{figure}
\autoref{fig:atct:phi} provides an example of a transposition base space change \dfunch, a scaling fiber space change \dfunct, and a composition of the two \dfunc\ applied to each data point $x_{\dbasepoint} \in \texttt{data}$. In the transposition only case, the values in $\dfuncpull\dsection$ retain their neighbors from $\dsection$ because \dfunc\ does not change the continuity. Each value in $\dfuncpull\dsection$ is also the same as in $\dsection$, just moved to the new position. In $\dfunct\dsection$, each value is scaled by two but remains in the same location as in $\dsection$. And in $\dfunc\dsection$ each function is transposed such that it retains its neighbors and all values are scaled consistently.

@@

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{equivariance.png}
  \caption{This artist is equivariant because when the input data $\dsection$ is transposed, $\dfunch$, scaled $\dfunct$, and transposed and scaled $\dfunc$, the corresponding colored cells are transposed, scaled such that the color is moved two steps, and both transposed and scaled. 
  \label{fig:artist:equivariance}}
\end{figure}

@@

\begin{equation*}
  \label{eq:artist:multiplication}
  \vartistc_{a \times b}(\cgamma{\dbase}{\dtotalc^{a\times b}}) \coloneqq \vartistc_{a}(\cgamma{\dbasec}{\dtotalc^{a}}) \times \vartistc_{b}(\cgamma{\dbasec}{\dtotalc^{b}}) 
\end{equation*}

which following from an adjoint property of homsets \note{find citation and push this into a footnote or appendix maybe} 
\begin{equation}
  Hom(\gbase, \gfiber) \times Hom(\gbase, \gfiber) = Hom(\gbase, \gfiber\times \gfiber)
\end{equation}

which means that the artists on the subsets of fibers can be defined 
\begin{equation}
  \gsectionc^{a \times b} = \{\gsectionc^{a}(\gbasepointc), \gsectionc^{b}(\gbasepointc)\}, \gbasepointc \in \vindexprec(\dbasec)
\end{equation} 
but that the signature of $\gsectionc^{a \times b}$ would be $\gbase \rightarrow \gfiber \times \gfiber$. Instead of having to special case the return type of artists that are compositions of multiple case, the hom adjoint \note{find cite} property
\begin{equation*}
  Hom(\gbase, \gfiber \times \gfiber) = Hom(\gbase+\gbase, \gfiber)
\end{equation*}
 means that multiplication can be considered as a special case of addition where $\dbase^{a} = \dbase^{b}$. While we discussed the trivial case in \autoref{sec:artist:addition}, there is no strict  requirement that $\dfiber^{a} = \dfiber^{b}$.

@@

\begin{figure}[H]
  \centering
  \includegraphics*[scale=.75]{qcom.png}
  \caption{The circle-line visual element can be constructed via $\gsection_{circle}$ + $\gsection_{line}$ functions that generate the circle and line elements respectively. This is equivalent to a $\gsection_{circle+line}$ function that takes as input the combined base space $\gbase_{circle} \sqcup \gbase_{line} = \gbase_{circle-line}$ and returns pixels in the circle-line element.  \label{fig:artist:operator}}
\end{figure}
There is no way to visually determine whether a visual element is the output of a single artist or a multiplied or added collection of artists. The circle-line visual element in \autoref{fig:artist:operator} can be a visual representation of a highlighted point intersecting with a line plot with the same fields. The same element can also be encoding some fields of a section in the circle and other fields of that section in the lines. \note{+*equive}
Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
\begin{equation}
  \label{eq:artist:join_base_fiber}
  \vartistc(\cgamma{\mathop{\sqcup}_{i} \dbasec^{i}}{\mathop{\oplus}_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
  \vartistc_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
\end{equation}

As shown in \autoref{eq:artist:join_base_fiber}, bundles over a union of base spaces can be joined as a product of the fibers. This allows us to consider all the data inputs in a complex visualization as a combined input, where some sections evaluate to null in fields for which there are no values for that point in the combined base space $\dbasepoint \in \mathop{\sqcup}_{i} \dbase^{i}$ The combined construction of the data is a method for expressing what each data input has in common with another data input-for example the data for labeling tick marks or legends- 
and therefore which commonalities need to be preserved in the artists that act on these inputs.

@@

\begin{figure}[H]

@@

\begin{figure}[H]
  \includegraphics[width=1\columnwidth]{full_scatter.png}
  \caption{This simple $\vmarkc$ assembles a circular visual element that is the color specified in $\vsection(\dbasepoint)$ and is at the intersection specified in $\vsection(\dbasepoint)$ \label{fig:construction:q} \note{much better labeling, include semantic labeling, make everything bigger}}
\end{figure}
As shown in \autoref{fig:construction:q}, a set of  $\vchannel$ functions individually convert the values in the data record to visual components. Then the $\vmark$ function combines these visual encodings to produce a graphic section \gsection. When this section is evaluated on the graphic space associated with the data $\gsection(\vindexpre(\dbasepoint))$, it produces a blue circular marker at the intersection of the x and y positions listed in \vsection. The composition rule in \autoref{eq:construction:q:fabrication} means that developers can implement $\vmark$ as drawing circles or can implement a $\vmark$ that draws arbitrary shapes, and then provide different $\vchannel$ adapters, such as one that specifies that the shape is a circle.

@@

\subsection{Bundle Inspired Data Containers}
\begin{table}[H]
  \centering
\begin{tabular}{|lcl|}
  \hline \\
   fruit &  calories &  juice \\
  \hline\\
    apple &        95 &   True \\ 
   orange &        67 &   True \\ 
  lemon &        17 &  False \\ 
      lime &        20 &  False \\
  \hline
\end{tabular}
\label{tab:discussion:data}
\end{table}
We construct a toy dataset with a discrete $\dbase$ of 4 points and a fiber space of $\dfiber=\{apple,\, orange,\, lemon\}\times \mathbb{Z}^{+} \times \{\texttt{True}, \texttt{False}\}$. We thinly wrap \autoref{tab:discussion:data} in an object so that the common data interface function is that $\tau = \texttt{DataContainerObject.query}$. 
\begin{minted}{python}
  class FruitFrameWrapper:
    def query(self, data_bounds, sampling_rate):
      # local sections are a list of
      # {field: local_batch_of_values}
      return local_sections
\end{minted}

This interface provides a uniform way of accessing subsets of the data, which are local sections. The motivation for a common data interface is that it would allow the artist to talk to different common python data containers, such as numpy\cite{harris2020array}, pandas\cite{jeff_reback_2020_3715232}, xarray \cite{hoyer2017xarray}, and networkx\cite{HagbergExploringNetwork2008}. Currently, data stored in these containers must be unpacked and converted into arrays and matrices in ways that either destroy or recreate the structure encoded in the container. For example a pandas data frame must be unpacked into its columns before it is sent into most artists and continuity is implicit in the columns being the same length rather than a tracked base space $\dbase$. Because it is more efficient to work with the data in column order, we often project the fiber down into individual components. As shown in \autoref{eq:atct:fiber_product}, we can verify that this projection is correct by checking that the values at the index are the same regardless of the level of decomposition.

@@

\begin{table}[H]
  \centering
  {\renewcommand{\arraystretch}{1.5}
  \begin{tabular}{|r | c c c|}
    \hline
    &\textcolor{base}{point}/\textcolor{base}{openset}/\textcolor{base}{base space} & \textcolor{fiber}{fiber space} & \textcolor{total}{total space}\\
     &  location/subset/indices & record/fields &  dataset type\\
    \hline
   Data & $\dbasepointc \in \opensetc \subseteq \dbasec$ & $\delementc \in \dfiberc$ & \dtotalc\\
   Visual & $\dbasepointc \in \opensetc \subseteq \dbasec$  & $\velementc \in \vfiberc$ & \vtotalc\\
   Graphic & $\gbasepointc \in \opensetgc \subseteq \gbasec$ & $\gelementc \in \gfiberc$ & \gtotalc\\
   \hline
  \end{tabular}
  \caption{Topological spaces introduced in \autoref{sec:atct:fiber-bundles}}
  \label{tab:appendix:summary:objects}
  }
\end{table}

\begin{table}[H]
  \centering
  {\renewcommand{\arraystretch}{1.5}
  \begin{tabular}{|r | l l | }
    \hline
     & \textcolor{section}{section} & \textcolor{sheaf}{sheaf} \\
     & record at location & set of possible records for subset \\
     \hline
  Data & $ \cgamma{\dbasec}{\dtotalc} \ni \dsectionc: \dbase \textcolor{section}{\rightarrow} \dfiberc$ & $\sheafc_{\dbasec, \dtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$\\
  Visual &  $\cgamma{\dbasec}{\vtotalc} \ni \vsectionc: \dbase \textcolor{section}{\rightarrow} \vfiberc$ & $\sheafc_{\dbasec, \vtotalc}: \opensetc \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}$\\
  Graphic &    $\cgamma{\gbasec}{\gtotalc} \ni \gsectionc: \gbase \textcolor{section}{\rightarrow} \gfiberc$ &  $\sheafc_{\gbasec, \gtotalc}: \opensetgc \rightarrow \cgamma{\opensetc}{\gtotalc\restriction_{\opensetgc}}$ \\ 
  \hline
  \end{tabular}
  \caption{Functions that associate topological subspaces with records, discussed in \autoref{sec:atct:fb:sections} and \autoref{sec:atct:sheaves}}
  \label{tab:appendix:summary:datafunctions}
  }
\end{table}

\begin{table}[H]
  \centering
  {\renewcommand{\arraystretch}{1.5}
  \begin{tabular}{|r | l  l |}
\hline
& function & constraint\\
\hline
 \gbasepointc\ to \dbasepointc& $ \vindexc: \opensetgc \rightarrow \opensetc$ & for $\gbasepointc \in \opensetgc$ exists $\dbasepointc \in \opensetc$ \\
 & & s.t. $\vindexc(\gbasepointc) = \dbasepointc$\\
 graphic for \dbasepointc & $ \vindexpushc \gsectionc: \opensetc \rightarrow \vindexpushc\gtotalc\restriction_{\opensetc}$ & $\vindexpushc\gsectionc(\dbasepointc)(\gbasepointc) = \gsectionc(\gbasepointc)$ \\
 record for \gbasepointc & $\vindexpullc\dsectionc: \opensetgc \rightarrow  \vindexpullc\dtotalc\restriction_{\opensetgc}$ & $\vindexpullc\dsectionc(\gbasepointc)=\dsectionc(\vindexc(\gbasepointc)) = \dsectionc(\dbasepointc)$  \\
\hline
  \end{tabular}
  \caption{Functors between graphic and data indexing spaces  \autoref{sec:atct:xi}}
  \label{tab:appendix:summary:transport}
  }
\end{table}
